$\pi$ with leftovers: \\ a mechanisation in Agda

### 2020-05-05 paper/paper.tex
on auxiliary results

\todo{intro}

\subsection{Auxiliaty Results}
\label{auxiliary-results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{nitheorem}[Framing]
  \label{thm:framing}
  The well-typedness of a process is independent of its leftover resources.
  Let $P$ be a well-typed process $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
  Let there be a $\Delta$ such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
  Take arbitrary contexts $\Gamma_r$ and $\Xi_r$ such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.
  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
\end{nitheorem}

\begin{nitheorem}[Weakening]
  \label{thm:weakening}
  Inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.

  Let $\func{insert}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then lifting every variable greater than or equal to $i$ in $P$ is well-typed under $\types{\func{insert}_i \; t \; \gamma}{\func{insert}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{insert}_i \; x \; \Xi}$.
\end{nitheorem}

\begin{nitheorem}[Strengthening]
  \label{thm:strengthening}
  Removing an unused variable preserves the well-typedness of a process.

  Let $\func{delete}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Let $i$ be a variable index not mentioned by $P$, such that $\Unused_i \; P$.
  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed under $\types{\func{delete}_i \; \gamma}{\func{delete}_i \; \Gamma}{\func{lower}_i \; P}{\func{delete}_i \; \Xi}$.
\end{nitheorem}

@@

\begin{nitheorem}[Swapping]
  \label{thm:swapping}
  Swapping two variables preserves the well-typedness of a process.
  Let $\func{swap}_i$ be a function that swaps variables $i$ and $\suc i$.
  For simplicity, we overload $\func{swap}$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
  \todo{}
  Let $P$ be a well-typed process such that $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
\end{nitheorem}

@@

\begin{proof}
  All the above theorems are by induction on \textsc{Process} and \textsc{Var}.
  For details, refer to the mechanisation in Agda \cite{Zalakain2020Agda}.
\end{proof}

>>>>

In \autoref{auxiliary-results} we present some auxiliary results: framing (\autoref{thm:framing}), weakening (\autoref{thm:weakening}), strengthening (\autoref{thm:strengthening}) and swapping of variables (\autoref{thm:swapping}).
We present subject congruence in \autoref{subject-congruence} and substitution in \autoref{substitution}.
Finally, we present subject reduction in \autoref{subject-reduction}.

\subsection{Auxiliaty Results}
\label{auxiliary-results}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
Framing asserts that the well-typedness of a process is independent of its leftover resources.
\begin{nitheorem}[Framing]
  \label{thm:framing}
  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma_l}{P}{\Xi_l}$.
  Let $\Delta$ be such that $\opctx{\Gamma_l}{\Delta}{\Xi_l}$.
  Let $\Gamma_r$ and $\Xi_r$ be arbitrary contexts such that $\opctx{\Gamma_r}{\Delta}{\Xi_r}$.

  Then $\types{\gamma}{\Gamma_r}{P}{\Xi_r}$.
\end{nitheorem}

Weakening states that inserting a new variable into the context preserves the well-typedness of a process as long as the usage annotation of the inserted variable is preserved as a leftover.
Let $\func{ins}_i$ insert an element into a context at position $i$ --- for simplicity, we use it both to insert types into typing contexts and usage annotations into usage contexts.
\begin{nitheorem}[Weakening]
  \label{thm:weakening}
  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, lifting every variable greater than or equal to $i$ in $P$ is well-typed in
  $\types{\func{ins}_i \; t \; \gamma}{\func{ins}_i \; x \; \Gamma}{\func{lift}_i \; P}{\func{ins}_i \; x \; \Xi}$.
\end{nitheorem}

Removing an unused variable preserves the well-typedness of a process.
Let $\func{delete}_i$ delete the element at position $i$ from a context --- for simplicity, we use it both to delete types from typing contexts and usage annotations from usage contexts.
\begin{nitheorem}[Strengthening]
  \label{thm:strengthening} 
  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Let $i$ be a variable index not in $P$, such that $\Unused_i \; P$.
  Then lowering every variable greater than or equal to $i$ in $P$ is well-typed in $\types{\func{delete}_i \; \gamma}{\func{delete}_i \; \Gamma}{\func{lower}_i \; P}{\func{delete}_i \; \Xi}$.
\end{nitheorem}

@@

Swapping two variables preserves the well-typedness of a process.
Let $\func{swap}_i$ be a function that swaps variables $i$ and $\suc i$.
For simplicity, we overload $\func{swap}$ and use it to swap types in a typing context, multiplicities in a usage context, and variable references in a process.
\begin{nitheorem}[Swapping]
  \label{thm:swapping}
  Let $P$ be a well-typed process in $\types{\gamma}{\Gamma}{P}{\Xi}$.
  Then, $\types{\func{swap}_i \; \gamma}{\func{swap}_i \; \Gamma}{\func{swap}_i \; P}{\func{swap}_i \; \Xi}$.
\end{nitheorem}

@@

\begin{proof}
  All the above theorems are proved by induction on \textsc{Process} and \textsc{Var}.
  For details, refer to the mechanisation in Agda \cite{Zalakain2020Agda}.
\end{proof}
